I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-07 paper_io_sim.tex
edits in methods

    \section{Method}
    \label{method}

        In this section, we present our page cache simulation model for file I/O,
        flusing, and eviction. We detail the design and implementation of two simulators:
        (1) a standalone Python prototype to evaluate the simulation model independently of any other 
        tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
        We also describe experimental scenarios to evaluate our simulators in various configurations.

        \subsection{Principle of the simulator}

            We separate our simulation model in two components, the IO
            Controller and the Memory Manager, which together simulate 
            file reads and writes in Linux (Fig.~\ref{fig:interaction}). 
            To read or write a file chunk, a simulated application sends a
            request to the IOController. The IOController interacts as needed with
            the Memory Manager to free memory through flushing or eviction,
            and to read or write to cache or to disk. The Memory Manager
            implements these operations, and simulates periodical flushing
            and eviction from cache.

            \begin{figure}
                   \centering
                   \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
                   \caption{Overview of the simulated writeback page cache.
                   Applications send file read or write requests to the
                   IOController that orchestrates flushing, eviction, cache
                   and disk accesses with the MemoryManager. Storage
                   devices (memory and disk) simulate file transfers with
                   bandwidth sharing. \tristan{Could you annotate arrows to
                   make them more explicit?}.}
                   \label{fig:interaction}
            \end{figure}

            \subsubsection{Memory simulation model}

            We use existing storage simulation models~\cite{lebre2015} to simulate memory,
            characterized by its storage capacity, read and write bandwidths,
            and latency. Reusing existing storage models allows us 
            to simulate bandwidth sharing between concurrent memory accesses.

            \paragraph*{Memory Manager}
            The Memory Manager simulates two parallel threads: the main one
            implements flushing, eviction, and cache I/Os synchronously,
            whereas the second one, operating in the background,
            periodically searches for expired dirty data in page cache LRU
            lists and flushes them to disk.

            \paragraph*{Page cache LRU lists}
            In the Linux kernel, page cache LRU lists contain file pages. However, 
            due to the large number of file pages, managing lists of pages 
            induces substantial overheads.
            Thus, we introduce the concept of a data block as a unit to represent data 
            cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
            A data block has information about file name, block size, last access 
            time, a dirty flag that represents whether the data is clean (0) 
            or dirty (1), and the time it is written to cache as dirty data. 
            Blocks can have different sizes and a given file can have multiple 
            data blocks in page cache. In addition, a data block can be split into an 
            arbitrary number of smaller blocks.

            \begin{figure}
                   \centering
                   \includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
                   \caption{Model of page cache LRU lists with data blocks.}    \label{fig:lrulist}
            \end{figure}

            We model page cache LRU lists as 
            two lists of data blocks, an active list and an inactive list, both ordered by 
            last access time (earliest first, Fig.~\ref{fig:lrulist}).
            As in the kernel, our simulator limits the size of the active list to
            twice the size of the inactive list, by moving least recently 
            used data blocks from the active list to the inactive list.
            Modeling page cache LRU lists as lists of data blocks reduces the
            overhead of the simulator, as data blocks are obviously less
            numerous than file pages, and at the same time achieves the accuracy in 
            I/O time simulation that can be seen in the later section.

            At a specific time, an arbitrary file can be partially cached, completely cached, 
            or uncached.
            Cached data is composed of data blocks existing in one or both
            of the page cache LRU lists. 
            Initially, blocks accessed for the first time are added to the inactive list in 
            page cache. 
            If the blocks residing the inactive list are subsequently accessed, 
            they will be moved to the top of the active list
            Cached blocks written to cache are marked as dirty blocks until 
            they are flushed to disk.

            \paragraph*{File read and write}
            Our model supports chunk-by-chunk file read and write simulation
            with a user-defined chunk size, in which chunks are read/written
            until file is entirely read/written.
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            In the scope of this study, our model support file accesses in chunk 
            but not file random access.
            Thus, we simplify the data read order, in which uncached data is read 
            from disk first, and cached data is only read after all uncached data is read 
            (left to right in Fig.~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put at the top of the inactive list. 
            For a chunk of cached data, one or more existing data blocks in two 
            page cache LRU lists are accessed again, and moved to the active list. 
            If the blocks are clean, we only need to keep the information about the 
            amount of data as they are available for eviction. 
            Thus, these re-accessed clean blocks can be merge into a bigger block 
            to reduce the number of block objects and put it in the active list. 
            If the blocks are dirty, since the blocks need to keep the information 
            of entry time for periodical flushing, they are only moved to the active list. 
            Because the chunk size and block size can be different, there are circumstances 
            in which a block is not entirely read. 
            In this case, the block is split into two smaller blocks, one is re-accessed 
            and the other remains the same.

            \begin{figure}
                   \centering
                   \includegraphics[width=\columnwidth]{figures/read_order.pdf}
                   \caption{File data read order from left to right, in which uncached data 
                   is read first, then cached data from inactive list and active list is read later.}    
                   \label{fig:read_order}
            \end{figure}

            For writes, we assume that all data to be written is 
            uncached and it is written to cache as dirty data. 
            Thus, each time a chunk is written, we create a block of dirty data 
            and put it in the inactive list.

            As is done in the kernel, we ensure the size of the active list does not 
            exceed twice the size of the inactive list 
            \cite{gorman2004understanding, linuxdev3rd2010}. 
            This is done by moving least recently used blocks in the active list 
            to the inactive list when the active list grows. If only a part of a data 
            block needs to be moved, the block is split.

            \paragraph*{Flushing}
            Next, we simulate data flushing, which runs in the main simulated 
            thread and writes back dirty data to disk. 
            The data flushing simulation function traverses the sorted 
            inactive list first, and then the active list, looks for least recently used 
            dirty blocks, writes them to disk through the storage device model, 
            and sets dirty flags of these blocks to 0 until the total flushed amount 
            reaches the amount to be flushed or there is no dirty data left in cache. 
            Similar to file read and write, the size of flushed data and block size 
            can be different, there are cases that the last block is not entirely flushed.  
            In that case, the last block is split into two blocks, one that is flushed 
            and one that remains dirty.
            The flushing time is simulated with storage model.

            \paragraph*{Cache eviction}    
            Similarly to data flushing, our cache eviction simulation runs on 
            the main thread, frees up the page cache by traversing and deleting 
            least recently used clean data blocks in the inactive list.
            The amount of evicted data is passed in and data blocks are deleted 
            from the inactive list until total evicted data reaches the amount 
            passed in, or there is no clean block left in the list.
            If the last evicted block is not entirely evicted, the block is split, 
            and only one block is deleted.
            The cache eviction simulation does not add up to the simulated time 
            since cache eviction time is negligible in real systems.

            \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
                \small
                \begin{algorithmic}[1]
                    \Input
                        \Desc{in}{page cache inactive list}
                        \Desc{ac}{page cache active list}
                        \Desc{t}{predefined flushing time interval}
                        \Desc{exp}{predefined expiration time}                        
                        \Desc{sm}{storage simulation model}
                       \EndInput
                       \While{host is on}
                        \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                        \State flushing\_time = sm.write(blocks)
                        \State flag blocks as clean data
                        \If{flushing\_time $<$ t}
                            \State sleep(t - flushing\_time)
                        \EndIf
                    \EndWhile
                \end{algorithmic}
            \end{algorithm}

            \paragraph*{Periodical flushing}    
            The periodical flushing simulation is run on a simulated background thread 
            to flush expired dirty data blocks from cache to disk.
            Similar to expired dirty pages in real Linux systems, a dirty block in our model 
            is considered expired if: 
            \begin{enumerate}
                \item It is dirty and 
                \item The time from its entry time to current time is longer than the 
                predefined expiration time. 
            \end{enumerate}
            The simulation algorithm of periodical flushing is an infinite loop in which 
            Memory Manager searches and flushes these expired dirty blocks to disk, 
            as described in Algorithm~\ref{alg:pdflush}. 
            The loop ends when the host is off (line 7). 
            In each repetition, Memory Manager finds expired dirty blocks in two 
            page cache LRU lists (line 8), simulates writes of data of these blocks 
            to disk (line 9), and mark them as clean (line 10).
            If the flushing time does not exceed our time interval, the thread is put 
            to sleep for the remaining time (lines 11-13). 
            Then based on the state of the host at current simulated time, 
            the algorithm continues or finishes the loop.
            Because periodical flushing is simulated as a background thread, it can be 
            concurrent with disk I/O on main thread. This parallelism is handled by the 
            storage model and reflected in simulated I/O time.

            \subsubsection{File I/O simulation model}

            As mentioned, in our model, file reads and writes can be completed by 
            reading or writing chunk-by-chunk until entire file is read or written.

            \begin{algorithm}\caption{File chunk read simulation of IOController}
            \label{alg:read}
                \small
                \begin{algorithmic}[1]
                    \Input
                        \Desc{cs}{chunk size}
                        \Desc{fn}{file name}
                        \Desc{fs}{file size (assume to fit in memory)}
                        \Desc{mm}{MemoryManager object}
                        \Desc{sm}{storage simulation model}
                       \EndInput
                       \State disk\_read = min(cs, fs - mm.cached(fn))
                       \State cache\_read = cs - disk\_read
                    \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
                    \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
                    \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                        \State sm.read(disk\_read)  
                        \State mm.add\_to\_cache(disk\_read, fn)     
                    \EndIf
                    \If {cache\_read $>$ 0} \Comment{Read cached}
                        \State mm.cache\_read(cache\_read)  
                    \EndIf
                \end{algorithmic}
            \end{algorithm}

            When a simulated application reads a file chunk, it sends a chunk read request 
            to IOController.             
            The algorithm to simulate a chunk read implemented in IOController 
            is described in Algorithm~\ref{alg:read}.   
            First, we calculate the amount of uncached data that needs to be read 
            from disk, and the remaining amount is read from cache (line 7-8).
            We assume that after a file is read, there are two copies of the file in memory 
            (one in cache, one in anonymous memory). 
            Unfortunately, SimGrid lacks of application memory management and WRENCH 
            only use the memory capacity to check if the applications can be run. 
            Thus, in our model, when a chunk is read, one copy of the chunk 
            occupies anonymous memory, and the uncached part of that chunk is 
            read into cache.    
            The amount of required memory to read the chunk is calculated and 
            if there is not enough memory available, Memory Manager is called to 
            flush dirty data to disk and simulate flushing time with memory model (line 9). 
            This flushing is complemented by eviction in the Memory Manager (line 10). 
            Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
            and this data of the file is added to cache (line 13).
            Finally, if cached data is read, Memory Manager is called to simulate a cache read  
            and update the corresponding data blocks in cache (line 16).

            \begin{algorithm}\caption{File chunk write simulation of IOController}
            \label{alg:write}
                \small
                \begin{algorithmic}[1]
                    \Input
                        \Desc{cs}{chunk size}
                        \Desc{fn}{file name}
                        \Desc{mm}{MemoryManager object}
                        \Desc{sm}{storage simulation model}
                       \EndInput
                    \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
                    \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                        \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                        \State mem\_amt = min(cs, mm.free\_mem)
                        \State mm.write(fn, mem\_amt) 
                    \EndIf
                    \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                        \State mm.flush(cs - mem\_amt)  
                        \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                        \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
                        \State sm.write(cs - mem\_amt)
                    \EndIf

                \end{algorithmic}
            \end{algorithm}

            Algorithm~\ref{alg:write} describes our simulation of chunk write in 
            the IOController. 
            In the chunk write, data can be written to cache with memory bandwidth 
            without waiting for flushing or being throttled before the amount of 
            dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
            Thus, our algorithm initially checks the remaining amount of data that 
            can be written as dirty data (line 6).
            If this amount is greater than 0, MemoryManager is asked to evict 
            data from cache to accommodate the written data (line 8). 
            If the amount to evict calculated and passed in the evict function is less than 0, 
            the eviction does nothing.
            After evicting cache, the actual amount of data that can be written to 
            page cache is calculated (line 9), a cache write is then simulated 
            with this amount through memory model and data written is added to 
            page cache LRU lists (line 10).
            After dirty\_ratio is reached, dirty data has to be flushed before new data 
            can be written to cache again. 
            If there is remaining chunk data that has not been written, we flush and 
            evict data from cache as much as possible so that the this data can be 
            written to cache (line 13-14). 
            Here, MemoryManager does not guarantee that the amount of flushed 
            and evicted data is enough for the write of remaining data. 
            In the case it is less than the required amount, the remaining data 
            being written to cache can be flushed and evicted right away. 
            Thus, the amount of remaining data written and kept in cache 
            after the write is limited by the amount free memory, and this amount 
            is added to cache (line 15). 
            Finally, the disk write is simulated with the remaining data amount (line 16), 
            and the chunk write algorithm finishes.

>>>>

    \section{Method}
    \label{method}
    In this section, we present our page cache simulation model for file I/O,
    flusing, and eviction. We detail the design and implementation of two simulators:
    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
    We also describe experimental scenarios to evaluate our simulators in various configurations.

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IOController. The IOController interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.
    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the simulated writeback page cache.
           Applications send file read or write requests to the
           IOController that orchestrates flushing, eviction, cache
           and disk accesses with the MemoryManager. Storage
           devices (memory and disk) simulate file transfers with
           bandwidth sharing. \tristan{Could you annotate arrows to
           make them more explicit?}.}
           \label{fig:interaction}
    \end{figure}

    \subsection{Memory Manager}

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cache I/Os synchronously, whereas
    the second one, operating in the background, periodically searches for
    expired dirty data in page cache LRU lists and flushes them to disk. We
    use existing storage simulation models~\cite{lebre2015} to simulate
    memory, characterized by its storage capacity, read and write
    bandwidths, and latency. Reusing existing storage models allows us to
    simulate bandwidth sharing between concurrent memory accesses.

    \subsubsection{Page cache LRU lists}

    In the Linux kernel, page cache LRU lists contain file pages. However, 
    due to the large number of file pages, managing lists of pages 
    induces substantial overheads.
    Therefore, we introduce the concept of a data block as a unit to represent data 
    cached in memory. A data block is a subset of file pages stored in
    page cache that were accessed in the same I/O operation. 
    A data block has information about file name, block size, last access 
    time, a dirty flag that represents whether the data is clean (0) 
    or dirty (1), and an entry (creation) time.
    Blocks can have different sizes and a given file can have multiple 
    data blocks in page cache. In addition, a data block can be split into an 
    arbitrary number of smaller blocks.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
           \caption{Model of page cache LRU lists with data blocks.}    \label{fig:lrulist}
    \end{figure}

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
    % Modeling page cache LRU lists as lists of data blocks reduces the
    % overhead of the simulator, as data blocks are obviously less
    % numerous than file pages. At the same time, it achieves the accuracy in 
    % I/O time simulation that can be seen in the later section.
    % From Tristan: I don't think the sentences above are necessary,
    % they are quite repetitive.

    At any given time, a file can be partially cached, completely cached,
    or uncached at all. Cached data blocks may reside in one or both of the
    page cache LRU lists. The first time they are accessed, blocks are
    added to the inactive list. On subsequent accesses, blocks of the
    inactive list are moved to \tristan{if they are moved then they can't
    be in both lists} the top of the active list. Cached blocks written to
    cache are marked as dirty until they are flushed to disk.

    \subsubsection{Reads and Writes}

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and put at the top of the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and put it in the active list. 
    If the blocks are dirty, we move them to the active list to preserve their entry time. 
    Because the chunk size and block size can be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list.}
           \label{fig:read_order}
    \end{figure}

    For file writes, we assume that all data to be written is 
    uncached. Thus, each time a chunk is written, we create a block of dirty data 
    and add it to the inactive list.

    \subsubsection{Flushing and eviction}

    \paragraph*{Flushing}
    The main simulated thread in the Memory Manager can flush the
    cache to write dirty data to disk. The data flushing simulation
    function takes the amount of data to flush as parameter. While
    the amount of data to flush is not reached and there are dirty
    blocks remaining in cache, this function traverses the sorted
    inactive list, then the sorted active list, and writes the
    least recently used dirty block to disk having set its dirty
    flag to 0. In case the amount of data to flush requires that a
    block be partially flushed, the block is split in two blocks,
    one that is flushed and one that remains dirty. Flushing time
    to disk is simulated with the storage model.

    \paragraph*{Cache eviction}
    Similarly to data flushing, our cache eviction simulation runs in
    the main thread. It frees up the page cache by traversing and deleting 
    least recently used clean data blocks in the inactive list.
    The amount of data to evict is passed as a parameter and data blocks are deleted 
    from the inactive list until total evicted data reaches the required amount,
    or until there is no clean block left in the list.
    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
    and only one of them is evicted.
    The cache eviction simulation does not contribute to the simulated time 
    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.

    \begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{in}{page cache inactive list}
                \Desc{ac}{page cache active list}
                \Desc{t}{predefined flushing time interval}
                \Desc{exp}{predefined expiration time}
                \Desc{sm}{storage simulation model}
               \EndInput
               \While{host is on}
                \State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
                \State flushing\_time = sm.write(blocks)
                \State flag blocks as clean data \tristan{a bit informal}
                \If{flushing\_time $<$ t}
                    \State sleep(t - flushing\_time)
                \EndIf \tristan{else?}
            \EndWhile
        \end{algorithmic}
    \end{algorithm}

    \paragraph*{Periodical flushing}
    Periodical flushing is simulated in the Memory Manager
    background thread. As in the Linux kernel, a dirty block
    in our model is considered expired if (1) it is dirty, and (2)
    the duration since its entry time is longer than a
    predefined expiration time. 
    Periodical flushing is simulated as an infinite loop in which 
    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
    % From Tristan: the algorithm is quite straightforward, I don't think this is 
    % necessary
    % In each repetition, Memory Manager finds expired dirty blocks in two 
    % page cache LRU lists (line 8), simulates writes of data of these blocks 
    % to disk (line 9), and mark them as clean (line 10).
    % If the flushing time does not exceed our time interval, the thread is put 
    % to sleep for the remaining time (lines 11-13). 
    % Then based on the state of the host at current simulated time, 
    % the algorithm continues or finishes the loop.
    Because periodical flushing is simulated as a background thread, it can happen concurrently
    with disk I/O initiated by the main thread. This parallelism is handled by the 
    storage model and reflected in simulated I/O time.

    \subsection{I/O Controller}

    \begin{algorithm}\caption{File chunk read simulation of IOController}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assume to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn))
               \State cache\_read = cs - disk\_read
            \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
            \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                \State sm.read(disk\_read)  
                \State mm.add\_to\_cache(disk\_read, fn)     
            \EndIf
            \If {cache\_read $>$ 0} \Comment{Read cached}
                \State mm.cache\_read(cache\_read)  
            \EndIf
        \end{algorithmic}
    \end{algorithm}            
    As mentioned previously, our model reads and writes file chunks in a round-robin 
    fashion.
    When a simulated application reads a file chunk, it sends a chunk read request 
    to IOController.             
    The algorithm to simulate a chunk read implemented in IOController 
    is described in Algorithm~\ref{alg:read}.   
    First, we calculate the amount of uncached data that needs to be read 
    from disk, and the remaining amount is read from cache (line 7-8).
    We assume that after a file is read, there are two copies of the file in memory 
    (one in cache, one in anonymous memory). 
    Unfortunately, SimGrid lacks of application memory management and WRENCH 
    only use the memory capacity to check if the applications can be run. 
    Thus, in our model, when a chunk is read, one copy of the chunk 
    occupies anonymous memory, and the uncached part of that chunk is 
    read into cache.    
    The amount of required memory to read the chunk is calculated and 
    if there is not enough memory available, Memory Manager is called to 
    flush dirty data to disk and simulate flushing time with memory model (line 9). 
    This flushing is complemented by eviction in the Memory Manager (line 10). 
    Then, if there is uncached data, a disk read is simulated with disk model (line 12), 
    and this data of the file is added to cache (line 13).
    Finally, if cached data is read, Memory Manager is called to simulate a cache read  
    and update the corresponding data blocks in cache (line 16).
    \begin{algorithm}\caption{File chunk write simulation of IOController}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write(fn, mem\_amt) 
            \EndIf
            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)  
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem))
                \State sm.write(cs - mem\_amt)
            \EndIf

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk write in 
    the IOController. 
    In the chunk write, data can be written to cache with memory bandwidth 
    without waiting for flushing or being throttled before the amount of 
    dirty data reaches dirty\_ratio, or otherwise written with disk bandwidth.
    Thus, our algorithm initially checks the remaining amount of data that 
    can be written as dirty data (line 6).
    If this amount is greater than 0, MemoryManager is asked to evict 
    data from cache to accommodate the written data (line 8). 
    If the amount to evict calculated and passed in the evict function is less than 0, 
    the eviction does nothing.
    After evicting cache, the actual amount of data that can be written to 
    page cache is calculated (line 9), a cache write is then simulated 
    with this amount through memory model and data written is added to 
    page cache LRU lists (line 10).
    After dirty\_ratio is reached, dirty data has to be flushed before new data 
    can be written to cache again. 
    If there is remaining chunk data that has not been written, we flush and 
    evict data from cache as much as possible so that the this data can be 
    written to cache (line 13-14). 
    Here, MemoryManager does not guarantee that the amount of flushed 
    and evicted data is enough for the write of remaining data. 
    In the case it is less than the required amount, the remaining data 
    being written to cache can be flushed and evicted right away. 
    Thus, the amount of remaining data written and kept in cache 
    after the write is limited by the amount free memory, and this amount 
    is added to cache (line 15). 
    Finally, the disk write is simulated with the remaining data amount (line 16), 
    and the chunk write algorithm finishes.
